### 2017-02-28 Fusion-cvpr2017-RebuttalKit/egrebuttal-fusion.tex
prepare for iccv

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}

\def\dxyred{\textcolor{red}}
\def\bE{\bf{E}}
\def\bL{\bf{L}}
\def\bX{\bf{X}}
\def\calL{\mathcal{L}}
\def\blambda{\mbox{{\boldmath $\lambda$}}}

\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{3651} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Late Fusion via Subspace Search with Consistency Preservation}

\maketitle
\thispagestyle{empty}

We thank all the reviewers.
%R3 is positive.
%R1 has some suggestions about technical analysis.
%R2 may have some misunderstandings about our paper.
We will take the comments seriously and will carefully revise the paper.
Code will be released.

\noindent\emph{\textbf{To Assigned Reviewer 1}}

{\bf Q1: The technical part in Sec 3.2 a little over-sold.}

{\bf A1:}
\textbf{Thanks for the comments.}
First, ALM is usually used for solving convex problems~[50] and (G. Liu etal, TPAMI 2013),
but our Problem~(2) is \textbf{non-convex},
and thus the convergence behavior of Algorithm 1 is unclear using standard ALM.
Second, we explicitly indicate the connection between the definition (Eq.8) and the computation (Algorithm 2) of Riemannian gradient by presenting Lemma 1,
which has never been discussed in previous works.
Third, we present every step of Algorithm~1 to make the paper self-contained.
To address the concerns, we will add more references (including G. Liu etal, TPAMI 2013) and explain the difference between our paper and the related works.

Typos will be corrected.

\noindent\emph{\textbf{To Assigned Reviewer 2}}

\textbf{Thanks for constructive comments and suggestions.}
We may have not clearly presented our method so that reviewer have misunderstood some points.
Below we mainly clarify reviewer's misunderstandings, illustrate the contributions and highlight the differences between the existing methods.

It it correct that our approach is more efficient than SVD,
but it seems that reviewer may have misunderstood the connections/differences between the STOC paper (P. Jain et al. 2013) and our approach.
At the first glance the STOC paper looks directly applicable to our problem,
but the conditions of the paper and ours are quite different.
Specifically, the STOC paper requires a \textbf{Incoherent Matrix Assumption} for matrix completion,
which is not guaranteed to converge the global solution in our work.
Therefore, we disagree with reviewer that alternating minimization would reach a global solution with a simpler approach.
In addition, as shown in [41], the Riemannian manifold method converges faster than the alternating minimization approach.
%indicating that our approach is more capable of dealing with high-dimensional large scale data.

Apart from the misunderstanding, reviewer's concerns are mainly about the contributions given [10], [50] and (Liu et al. PAMI 2013).
Below, we brief the novelties and contribution over [10], [50] and (Liu et al. PAMI 2013).
%[10] and [50] formulate the fusion problem as a nuclear norm regularization problem, which cannot handle large-scale problems.
First, our method leverages $l_{1,2}$ norm for malicious classifier detection rather than the commonly used robust $l_{1}$ norm [10].
Second, we show the convergence of ALM for solving a non-convex problem rather than the convex problem in previous work (Liu et al. 2013).
Third, we propose to use matrix factorization (MF) technique to improve the efficiency comparing with nuclear norm based methods [10, 50].
Fourth, most MF methods only consider smooth $l_{2}$ loss, but we extend it to non-smooth $l_{1,2}$ loss.

Based on our discussions above,
Algorithm 2/3 should not be replaced by other methods based on the STOC paper.

Regarding the comments on Theorem 1, we clarify a few minor misunderstandings and technical issues.
First, the convergence of $\bX$ is guaranteed by [41].
Second, the proof of Theorem 1 relies on the gradient without the convexity requirement.
Third, the assumption for the boundedness of $\lambda_k$ is common (Th. Oh etal. TPAMI2016).
With $k \rightarrow \infty$, the gradient of $\calL$ vanishes.
Thus our algorithm converges and accumulation points will be achieved.

Thanks for the comments on empirical study.
However, reviewer may have overlooked important results from the tables.
Our method significantly outperforms LPBosst, as reported in Table 1 and 5.
Specifically, our method achieves much higher accuracy \textbf{(89.03\%)} than LPBoost \textbf{(86.24\%)} on UCF-101,
and CIFAR-100 \textbf{(77.72\% vs 76.59\%)}.

As indicated in Sec 4.4, we have provided very detailed approach for post-processing.
All the parameters of all algorithms are tuned by cross-validation with the same setting.
No human intervention was involved.
Therefore, the results should be reproducible and reliable.

It is the convention of CV papers to compare the results reported in recent papers on a standard setting.
CV researchers sometimes do not release their best model for business reasons,
but often share the not-that-new models with the community.
We show that combining a few weak models using our method,
the performance can be better than a single strong model.
To be more specific, the best single model on CIFAR-10 was reported in 12/2016 (see arXiv:1611.05431) with a test error of 3.58.
Our method fuse multiple single models which all perform worse than (arXiv:1611.05431), but can achieve low test error on CIFAR-10 (3.47 in Table 5).
This demonstrates the benefit of a good late fusion method.

We will take reviewer's comments seriously and will revise our paper carefully according to the comments.
We hope reviewer will find our paper contains novel ideas and acceptable for CVPR based on our explanations.

\noindent\emph{\textbf{To Assigned Reviewer 3}}

\textbf{Thanks for comments.}
Inappropriate figures, notations and typos will be revised/corrected.
$\ell_{1,2}$ loss preserves the fidelity within columns (lines 106-110),
and we will add more discussions and results of using $\ell_{2}$ norm.
Regarding the question on adversarial classifiers, we have discussed this issue in Sec4.1 \textbf{visualized in Fig.2.}
We will add more discussions and cite the related paper.
Thanks for the last suggestion, and it will be investigated in our future work.

\end{document}