### 2017-02-28 back-aaai/RLF_MF.tex

\def\year{2017}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm,mathrsfs,amsfonts,dsfont}
\usepackage{algorithm}
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{threeparttable}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{color}
\usepackage{epstopdf}
\usepackage{bm}
%\usepackage[vlined,boxed,ruled]{algorithm2e}
\usepackage{url}
\usepackage{xspace}

\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}

\usepackage{aaai17}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\def\M{\mathcal{M}}
\def\R{{\mathbb R}}
\def\bR{{\bf R}}
\def\U{{\bf U}}
\def\V{{\bf V}}
\def\diag{\mbox{diag}}
\def\bsigma{\mbox{{\boldmath $\sigma$}}}
\def\trsp{{\sf T}}
\def\I{{\bf I}}
\def\0{{\bf 0}}
\def\E{{\bf E}}
\def\G{{\bf G}}
\def\grad{{\text{grad}}}
\def\bZ{{\bf Z}}
\def\bfeta{\mbox{{\boldmath $\eta$}}}
\def\mT{{\mathcal T}}

\def\bA{{\bf A}}
\def\bB{{\bf B}}
\def\bN{{\bf N}}
\def\bM{{\bf M}}
\def\bD{{\bf D}}
\def\bI{{\bf I}}
\def\bE{{\bf E}}
\def\blambda{{\bm \lambda}}
\def\calL{{\mathcal{L}}}
\def\calC{{\mathcal{C}}}
\def\calS{{\mathcal{S}}}
\def\calF{{\mathcal{F}}}
\def\bL{{\bf L}}
\def\bO{{\bf O}}
\def\bU{{\bf U}}
\def\bV{{\bf V}}
\def\dsR{\mathds{R}}
\def\bX{{\bf X}}
\def\bx{{\bf x}}
\def\btx{{\tilde{\bf x}}}
\def\by{{\bf y}}
\def\bw{{\bf w}}
\def\btw{{\tilde{\bf w}}}
\def\tK{\tilde{K}}
\def\txi{\tilde{\xi}}
\def\tildeb{{\tilde{b}}}
\def\tphi{{\tilde{\phi}}}
\def\bz{{\bf z}}
\def\br{{\bf r}}
\def\bv{{\bf v}}
\def\bb{{\bf b}}
\def\bp{{\bf p}}
\def\bA{{\bf A}}
\def\bI{{\bf I}}

\def\bx{{\bf x}}
\def\bX{{\bf X}}
\def\by{{\bf y}}
\def\bY{{\bf Y}}
\def\bw{{\bf w}}
\def\bW{{\bf W}}
\def\balpha{{\bm \alpha}}
\def\bmu{{\bm \mu}}
\def\bK{{\bf K}}
\def\bg{{\bf g}}
\def\bp{{\bf p}}
\def\p{p}
\def\bP{{\bf P}}

\def\balpha{{\bm \alpha}}
\def\bbeta{{\bm \beta}}
\def\eg{{\emph{e.g.}}}
\def\zerocolumn{{\bf 0}}

\def\ttTP{{\tt TP}}
\def\ttFP{{\tt FP}}
\def\ttFN{{\tt FN}}

\def\st{{\text{s.t.}}}
\def\etc{\emph{etc}}
\def\wrt{\emph{w.r.t}}
\def\ie{\emph{ie}}
\def\eg{\emph{eg}}
\def\rank{{\text{rank}}}
\def\Tr{{\text{Tr}}}

\def\yanred{\textcolor{red}}
\def\yanblue{\textcolor{blue}}

\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Insert Your Title Here)
/Author (Put All Your Authors Here, Separated by Commas)}
\setcounter{secnumdepth}{0}
 \begin{document}

% The file aaai.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

%\title{Consistent Late Fusion for Visual Recognition}
\title{Late Fusion via Subspace Search with Consistency Preservation}
%% Efficient

\author{AAAI Press\\
Association for the Advancement of Artificial Intelligence\\
2275 East Bayshore Road, Suite 160\\
Palo Alto, California 94303\\
}

\maketitle

\begin{abstract}
% First: A big picture of the scenario which we would consider
In many real-world applications, data can be represented in multiple ways or multi-view features,
in order to describe various characteristics of data.
In this way, it could often improve the prediction performance by taking advantage of these feature together.
% Second: what BIG problem would we like to solve?
Late fusion, which combines the predictions of multiple features, is a commonly used approach to make the final decision for a test instance.
% Third: specify the detailed problem: find a precise position for our paper
However, it is ubiquitous that different features would dispute on the data, which results in performance degeneration.
So there is a requirement of outlier rejection for late fusion algorithms.
% Fourth: general introduction of our method
In this paper, we propose an efficient matrix factorization based approach to fuse predictions from multiple sources,
which could avoid the performance degeneration caused by the controversy of multiple features on test data.
% Some unvalued sentence(s)
Extensive experiments demonstrate that the proposed method is effective to remove outliers and improve fusion performance.

\end{abstract}

\section{Introduction}

%背景 多model融合
Feature representation of objects, which transforms raw data into numerical features, is a prerequisite for most real-world applications.
There are often multiple ways to generate numerical features from raw data for visual recognition.
For example, for image data, one can construct hand-crafted features such as scale-invariant feature transform
(SIFT) feature~\cite{loweijcv2004distinctive} and Histograms of Gradients (HOG) feature~\cite{dalalcvpr2005histograms},
or extract features based on well-trained convolutional neural networks (CNNs)~\cite{krizhevsky2012imagenet}.
Moreover, text data can be represented by term frequency–inverse document frequency (tf-idf)~\cite{manning2008introduction} and word2vec~\cite{mikoloviclr2013efficient}.
Each category of features tends to capture some specific characteristics of data,
such as textures of images and frequencies of text data.
When building a prediction model, the model trained on one kind of feautures could be biased and hence may incur unsatisfactory performance.
Existing works have investigated the performance improvement by combining multiple kinds of features in real-world applications~\cite{gehler2009feature,ye2012robust,xuiccv2013feature,lai2015learning}.

%% late fusion is a common sigle model -> single feature, 介绍late fusion
Late fusion, a typical approach to obtaining the final decision on testing data, aims to fuse the predictions from different features.
A commonly used method is to learn weights for predictions from various features to combine them, which has a number of variations~\cite{gehler2009feature,xuiccv2013feature,lai2015learning}.
Among these algorithms, multiple kernel learning (MKL)~\cite{lanckriet2004learning,Rakotomamonjy2008Simplemkl} is often used to train such a weighted combination.
However, a single feature may produce corrupted predictions on some data, which likely degenerates performance.
To deal with this issue, therefore, some works formulate it as a robust low rank matrix recovery problem~\cite{gaoijcai2016robust,ye2012robust} by nuclear norm minimization.
Nevertheless, nuclear norm minimization applied in these works requires singular value decompositions (SVDs), which would be computationally unaffordable for large matrices.

To prevent from the expensive nuclear norm minimization, in this paper, we propose a more efficient matrix factorization algorithm to recover the underlying low rank label matrix.
As for outlier detection, as shown in Figure~\ref{fig:framework} (notations are introduced shortly), we apply robust $\ell_{1,2}$ loss to remove the outlier columns in the label matrix.
Most existing matrix factorization optimization methods, however, only consider smooth loss functions, rather than the nonsmooth loss.
We thus in particular present an approach based on augmented Lagrangian multiplier (ALM) to extend matrix factorization methods to nonsmooth problems.
The main contributions are summarized as below:
\begin{itemize}
  %\item We propose an efficient late fusion algorithm based on matrix factorization.
  \item We formulate the late fusion problem as a fixed-rank robust matrix recovery problem based on matrix factorization, which intuitively preserves the consistency of predictions from different sources.
  \item We propose an ALM based algorithm to extend existing matrix factorization algorithms from smooth least square loss to nonsmooth $\ell_{1,2}$ loss.
  \item Empirical studies show significantly improvement of our proposed method compared to existing late fusion algorithms.
\end{itemize}

\section{Related Work}
%% Matrix factorization
%% Fusion Methods

MKL based method is a simple but efficient way to combine different features for multi-model fusion,
and many improved algorithms have been proposed by introducing more constraint, boosting methods and other optimization.
In~\cite{vedaldi2009multiple}, the authors use MKL to learn an optimal combination of exponential ${\chi}^2$ kernels capturing a different feature channel.
In~\cite{gehler2009feature}, boosting approaches are proposed to ensemble different features, which learns the correct weighting of multiple predicted confidence scores from different models.
In~\cite{natarajan2012multimodal}, the authors propose a two-step strategy employing MKL and late score fusion methods to combine diverse features, specifically on event detection tasks.
In~\cite{lai2015learning}, a sample specific late fusion based on graph Laplacian with RankSVM style constraints is proposed,
which learns sample specific fusion weights by diffusing the information from labeled samples to the unlabeled one.
In~\cite{xuiccv2013feature}, a optimization algorithm are proposed to combine diverse decision values,
which learns the weights, thresholding and smoothing parameters in a joint framework.
However, all these mkl based fusion methods or others only learn weights to combine different basic models, which do not consider the outlier rejection in the fusion process specifically.

%Our proposed late fusion algorithm is based on matrix factorization,
%which has attracted increasing attention recently.
Matrix factorization has attracted increasing attention recently.
Many manifold optimization algorithms on matrix factorization have been proposed by exploiting the manifold geometry.
Based on manifold optimization, fixed-rank matrices are supposed to belong to a smooth matrix manifold~\cite{Absil2008OAMM,vandereycken2013lowrank}.
In~\cite{vandereycken2013lowrank}, the authors propose a low rank geometric conjugate gradient (LRGeomCG) method.
In~\cite{rtrmc2011boumal}, first- and second-order Riemannian trust-region methods are applied to solve the low rank matrix completion by exploiting the low rank constraint.
In~\cite{Bonnabel2011}, a linear regression algorithm is proposed, where the parameter of this algorithm is a fixed-rank matrix based on the Riemannian manifold geometry.
In~\cite{Mishra2012}, the authors propose a quotient geometric matrix completion method.
The authors in~\cite{grouse2010Balzano} propose an online algorithm for tracking subspaces, Grassmannian rank-one update subspace estimation.
In~\cite{ngonips2012scaled}, a gradient methods based on scaled metric on the Grassmann manifold is proposed.
The authors in~\cite{Wen2012} propose a low-rank matrix fitting algorithm is proposed to solve large scale matrix completion problems by exploiting nonlinear successive over-relaxation.
In~\cite{RechtNIPS2011hogwild}, a lock-free approach to parallelizing stochastic gradient descent is proposed.
However, all these matrix factorization methods only consider the general least square loss, which cannot be directly applied to handle $\ell_{1,2}$ loss.

%% Latent Fusion,

\section{The Proposed Approach}

%We first elaborate the formulation of our proposed algorithm.
We first elaborate the formulation of our proposed algorithm,
which regards the late fusion problem as a robust matrix recovery problem.
%Then we show the details about how to use augmented Lagrangian multiplier (ALM) to solve the $\ell_{1,2}$ objective function.
And then we present the details of using ALM extending matrix factorization to handle $\ell_{1,2}$ loss.
Finally we introduce our post-process strategy to convert the recovered label matrix to final predictions.

\begin{figure}[ht]
\centering\includegraphics[width=0.45\textwidth]{resource/frame_work.pdf}
\caption{Framework (example of $n=10, \calC=3, m=3$)}
\label{fig:framework}
\end{figure}

\subsection{Problem Formulation}

Suppose there are $n$ testing instances and $\calC$ classes.
We denote each instance as $x_i \in \dsR^{d}$ where $1 \leq i \leq n$.
Assume that we have already trained $m$ classifiers, \ie. $f_1, f_2, ... f_m$.
Then by applying each single classifier $f_j$ on the entire testing data, one can derive an indicator matrix $\bL_j \in \dsR^{n \times m}$, where $\bL_{j, ic} = 1$ if the $i$-th instance is assigned to the $c$ class, otherwise $\bL_{j, ic} = 0$ (as shown in Figure~\ref{fig:framework}).
Let $\bL = [ \bL_1, \bL_2, ..., \bL_\calC ]$.
%%On one hand, we aim to preserve the consistency among $m$ classifiers, so we introduce a rank-$m$ constraint on $\bL$.
On one hand, we aim to preserve the consistency among $m$ classifiers, so we introduce a rank constraint on $\bL$.
On the other hand, to catch the outlier columns, we apply robust $\ell_{1, 2}$ loss, which is computed by
{\small
$$ || \bX ||_{1,2} = \sum_{j = 1}^{m\calC} \sqrt{ \sum_{i = 1}^{n} \bX_{ij}^2 } . $$
}
\noindent
Thus, the proposed model can be written as below:
{\small
\begin{align}\label{eq:l12_rankconstraint}
  \min_{\bX} ~&~ || \bL - \bX ||_{1,2}    \nonumber \\
  \st ~&~ \rank(\bX) = p   ,
\end{align}
}
\noindent
where $\bX \in \dsR^{n \times m\calC}$ is the underlying low rank label matrix to be recovered, $\rank(\bX)$ obtains the rank of $\bX$,
and $p$ is a scalar.
In our setting, to preserve the consistency among various classifiers, we set $p = \calC$.

It is difficult to optimize Problem~(\ref{eq:l12_rankconstraint}) directly due to the presence of the rank constraint.
We hence transform the original problem as a matrix factorization formulation as below:
{\small
\begin{align}\label{eq:l12_mf}
  \min_{\bX} ~&~ || \bL - \bX ||_{1,2}   \nonumber \\
         \st ~&~ \bX = \bU \bV,~\bU \in \dsR^{n \times p},~\bV \in \dsR^{p \times m\calC}  .
\end{align}
}
\noindent
Nevertheless, existing matrix factorization algorithms focus on smooth loss functions, rather than a nonsmooth $\ell_{1,2}$ loss function~\cite{tanicml2014riemannian,vandereycken2013lowrank,Wen2012,ngonips2012scaled,rtrmc2011boumal}.
In the following section, we present an ALM optimization framework to solve Problem~(\ref{eq:l12_mf}).

%\begin{small}
\begin{algorithm}[ht]

\begin{algorithmic}

\REQUIRE $\bL \in \dsR^{n \times m\calC}$, rank($\bX$) = $p$

\STATE Initialize $\rho > 1$, $t = 0$, $\blambda_{0} = 0$, $\bE_{0} = 0$, and $\mu_{0} > 0$.

\WHILE{not converge}

  \WHILE{not converge}

    %\STATE 1: Obtain $\bX^{(t)}$ by solving Problem~(\ref{eq:subproblem_wrt_X}) via LRGeomCG.
    \STATE 1: Obtain $\bX_{t+1}$ by solving Problem~(\ref{eq:subproblem_wrt_X}).

    %\STATE 2: Obtain $\bE^{(t)}$ by column-wise soft-thresholding~(\ref{eq:column_wise_soft_thresholding}).
    \STATE 2: Obtain $\bE_{t+1}$ by solving Problem~(\ref{eq:column_wise_soft_thresholding}).

  \ENDWHILE

  \STATE 3: $\blambda_{t+1} = \blambda_{t} + \mu_{t} (\bL - \bX_{t+1} - \bE_{t+1})$.

  \STATE 4: $\mu_{t+1} = \rho \mu_{t}$.

  \STATE 5: t = t + 1.

\ENDWHILE

\ENSURE $\bX \in \dsR^{n \times m\calC}$

\end{algorithmic}
\caption{The ALM algorithm for Problem~(\ref{eq:mf_l21_constrained})}
\label{alg:alm_mf}
\end{algorithm}
%\end{small}

%
%\begin{small}
%\begin{algorithm}[h!]
%
%\begin{algorithmic}
%
%\REQUIRE $\bL \in \dsR^{n \times m\calC}$, $p$ (the rank of $\bX$)
%\ENSURE $\bX \in \dsR^{n \times m\calC}$
%\STATE Initialize $\rho > 1$, $t = 0$, $\blambda^{(0)} = 0$, and $\mu^{(0)} > 0$.
%
%\WHILE{not converge}
%
%  \WHILE{not converge}
%
%    %\STATE 1: Obtain $\bX^{(t)}$ by solving Problem~(\ref{eq:subproblem_wrt_X}) via LRGeomCG.
%    \STATE 1: Obtain $\bX^{(t)}$ by solving Problem~(\ref{eq:subproblem_wrt_X}).
%
%    %\STATE 2: Obtain $\bE^{(t)}$ by column-wise soft-thresholding~(\ref{eq:column_wise_soft_thresholding}).
%    \STATE 2: Obtain $\bE^{(t)}$ by solving Problem~(\ref{eq:column_wise_soft_thresholding}).
%
%  \ENDWHILE
%
%  \STATE 3: $\blambda^{(t+1)} = \blambda^{(t)} + \mu^{(t)} (\bL - \bX^{(t+1)} - \bE^{(t+1)})$.
%
%  \STATE 4: $\mu^{(t+1)} = \rho \mu^{(t)}$.
%
%  \STATE 5: t = t + 1.
%
%\ENDWHILE
%
%\end{algorithmic}
%\caption{The ALM algorithm for Problem~(\ref{eq:mf_l21_constrained})}
%\label{alg:alm_mf}
%\end{algorithm}
%\end{small}

\subsection{Optimization Based on ALM}

Most existing matrix factorization optimization methods only consider smooth loss functions, rather than the nonsmooth $\ell_{1,2}$ loss used in Problem~(\ref{eq:l12_mf}).
In this section, we present an ALM based algorithm to optimize Problem~(\ref{eq:l12_mf}).

By introducing a new variable $\bE = \bL - \bX$, we can develop Problem~(\ref{eq:l12_mf}) as below:
{\small
\begin{align}\label{eq:mf_l21_constrained}
  \min_{\bE, \bX} ~&~ || \bE ||_{1,2}      \nonumber \\
  \st             ~&~ \bE = \bL - \bX  ,   \nonumber \\
                  ~&~ \bX = \bU \bV,~\bU \in \dsR^{n \times p},~\bV \in \dsR^{p \times m \calC}
\end{align}
}
\noindent
The augmented Lagrangian function of Problem~(\ref{eq:l12_mf}) is constructed as below
{\small
\begin{align}\label{eq:lagrangian_l21}
  \calL (\bX, \bE, \blambda, \mu) = &~ || \bE ||_{1,2} + \langle \blambda, \bL - \bX - \bE \rangle      \nonumber \\
                                    & + \frac{\mu}{2} || \bL - \bX - \bE ||_F^2
\end{align}
}
\noindent
where $\bX = \bU \bV$, $\mu$ is a scalar,
and $\blambda \in \dsR^{n \times m\calC}$ are the Lagrangian multipliers.
Here $\langle \bA, \bB \rangle$ is the inner product of $\bA \in \dsR^{m \times n}$ and $\bB \in \dsR^{m \times n}$, and defined as $\langle \bA, \bB \rangle = \sum_{i=1}^{n} \sum_{j=1}^{m} \bA_{ij} \bB_{ij}$.
Then we can update $\bX$ and $\bE$ alternatively, which is summarized in Algorithm~\ref{alg:alm_mf}.

\begin{theorem}
\label{theorem:alm_convergence}
  Suppose that the sequence $\{ \blambda_{k} \}$ is bounded, any accumulation point $(\bX^*, \bE^*)$ of Algorithm~\ref{alg:alm_mf} is a stationary point.
\end{theorem}

\begin{proof}
\label{proof:proof_AA}
  The proof can be found in Appendix A.
\end{proof}

\subsection{Optimization of The Subproblem of $\bX$}

Particularly, to update $\bX$, we have the following augmented Lagrangian function \wrt. $\bX$:
{\small
\begin{align}\label{eq:langrangian_wrt_X}
  \calL (\bX) = & \langle \blambda, \bL - \bX - \bE \rangle + \frac{\mu}{2} || \bL - \bX - \bE ||_F^2  \nonumber  \\
              %= & \frac{\mu}{2} \Big( || \bL - \bX - \bE ||_F^2 + \frac{2}{\mu} \langle \blambda, \bL - \bX - \bE \rangle + \frac{|| \blambda ||_F^2}{\mu^2} \Big)    \nonumber   \\
              %  & - \frac{\mu}{2} \frac{|| \blambda ||_F^2}{\mu^2}   \nonumber \\
              %= & || \bL - \bX - \bE + \frac{\blambda}{\mu} ||_F^2 - \frac{1}{2\mu}|| \blambda ||_F^2   \nonumber \\
              = & || \bM - \bX ||_F^2 - \frac{1}{2\mu}|| \blambda ||_F^2   ,
\end{align}
}
\noindent
where $\bM = \bL - \bE + \frac{\blambda}{\mu}$.
Hence, we derive the following subproblem \wrt. $\bX$:
{\small
\begin{align}\label{eq:subproblem_wrt_X}
  \min_{\bX} ~&~ || \bM - \bX ||_F^2    \nonumber \\
  \st        ~&~ \bX = \bU \bV,~\bU \in \dsR^{n \times p},~\bV \in \dsR^{p \times m\calC}   .
\end{align}
}
\noindent
%Problem~(\ref{eq:subproblem_wrt_X}) contains a smooth loss function \wrt. $\bX$, thus can be efficiently solved by existing fixed-rank matrix factorization optimization algorithms, such as LRGeomCG~\cite{vandereycken2013lowrank}.
Problem~(\ref{eq:subproblem_wrt_X}) contains a smooth loss function of $\bX$.
We thus propose to solve the subproblem by exploiting Riemannian geometry of smooth fixed-rank matrices, which is briefly presented below.

The key idea of manifold optimization is that update of $\bX$ is always performed on the same manifold, which ensures the rank of $\bX$ unchanged as the optimization iterates.
Let $r$ be a known scalar.
A smooth manifold of fixed-rank-$r$ matrices is defined as~\cite{vandereycken2013lowrank}:
{\small
\begin{eqnarray}
\M_{r} &=& \{\bX\in \R^{n\times m\calC}: \rank(\bX) = r\} \nonumber \\
       &=& \{\U\diag(\bsigma)\V^{\trsp}: \U \in \textrm{St}_{r}^{n}, \V \in
           \textrm{St}_{r}^{m\calC}, ||\bsigma||_0 = r\} \nonumber
\end{eqnarray}
}
\noindent
where $\textrm{St}_{r}^{n} = \{\U \in \R^{n\times r}:
\U^{\trsp}\U = \I \}$ dentoes the Stiefel manifold of $n\times r$ real and orthonormal matrices.
We denote the tangent space by $T_{\bX}\M_{r}$ of $\M_{r}$ at $\bX = \U\diag(\bsigma)\V^{\trsp} \in \R^{n \times m\calC}$,
which is obtained as below
{\small
\begin{eqnarray}\label{eq:tangent_vector}
&&T_{\bX}\M_{r} =  \{\U\bM\V^{\trsp} \!\!+\! \U_p\V^{\trsp} \!\!+\!
\U\V_p^{\trsp}: \bM \in \R^{r\times r}, \nonumber\\ && \U_p \in \R^{n\times r},
\U_p^{\trsp}\U = \0, \V_p \in \R^{m\calC \times r}, \V_p^{\trsp}\V = \0\}.
\end{eqnarray}
}
\noindent
One can define a metric $g_{\bX}({\bA},{\bB}) = \langle{\bA},{\bB}\rangle$ on $\M_{r}$,
with $\bX \in \M_{r}$ and ${\bA},{\bB} \in T_{\bX}\M_{r}$,
then $\M_{r}$ becomes a Riemannian manifold by restricting
$\langle\bA,\bB\rangle$
to the \emph{tangent bundle}, defined as the disjoint union of all tangent spaces:
$T\M_{r} = \bigcup_{\bX\in \M_{r}}\{\bX\} \times T_{\bX}\M_{r}
         = \! \{(\bX,\bP)\in \R^{n \times m\calC} \times \R^{n \times m\calC}: \bX \in \M_{r}, \bP \in T_{\bX}\M_{r}\}.$

Let $f(\bX) = || \bM - \bX ||_F^2$.
%Suppose that $\G$ is the gradient of the smooth function $f(\bX)$ in Euclidean space at $\bX = \U\diag(\bsigma)\V^{\trsp} $.
Suppose that $\G = \nabla f(\bX)$ in Euclidean space at $\bX = \U\diag(\bsigma)\V^{\trsp}$.
Then, according to~\cite{vandereycken2013lowrank}, the Riemannian gradient of $f(\bX)$ on $\M_r$ is given as the orthogonal
projection of $\G$ onto the tangent space at $\bX$:
{\small
\begin{eqnarray}\label{eq:grad}
\grad{f(\bX)} = P_{T_{\bX}\M_{r}}(\G)  ,
\end{eqnarray}
}
\noindent
where
$P_{T_{\bX}\M_{r}}(\bZ): \bZ \mapsto P_{U}\bZ P_{V} + P_{U}^{\perp} \bZ P_{V} + P_{U} \bZ P_{V}^{\perp}$
denotes the orthogonal projection of any
$\bZ \in \R^{n \times m\calC}$ onto the tangent space at $\bX = \U\diag(\bsigma)\V^{\trsp}$.
Here $P_U = \U \U^{\trsp}$ and $P_U^{\perp} = \I - \U \U^{\trsp}$ for any $\U \in \textrm{St}_{r}^{n}$.

%\begin{small}
\begin{algorithm}
  \begin{algorithmic}
    \REQUIRE $\bX = \U \diag(\sigma) \V^{\trsp} \in \M_{r}$, $\G$.

    \STATE 1. $\bR_{u} \leftarrow \G^{\trsp} \U$, $\bR_{v} \leftarrow \G \V$.

    \STATE 2. $\bM \leftarrow \U^{\trsp} \bR_{v}$.

    \STATE 3. $\U_{p} \leftarrow \bR_{v} - \U \bM$, $\V_{p} \leftarrow \bR_{u} - \V \bM^{\trsp}$.

    \ENSURE $\grad f(\bX) = \U \bM \V^{\trsp} + \U_{p} \V^{\trsp} + \U \V_{p}^{\trsp} \in T_{\bX}\M_{r}$.

  \end{algorithmic}
  \caption{Computation of $\grad f(\bX)$ (Algorithm 2 in~\cite{vandereycken2013lowrank})}
  \label{alg:compute_riemannian_grad}
\end{algorithm}
%\end{small}

\begin{lemma}
  Assuming that $\U_{p}$, $\V_{p}$ and $\bM$ are computed by Algorithm~\ref{alg:compute_riemannian_grad},
  we have ${\textnormal{grad}} f(\bX) = \U \bM \V + \U_{p} \V^{\trsp} + \U \V^{\trsp}_{p}$.
\end{lemma}

\begin{proof}
  The proof can be found in Appendix B.
\end{proof}

%\begin{small}
\begin{algorithm}
  \begin{algorithmic}
    \REQUIRE Initial $\bX_{1} \in \M_{r}$, tangent vector $\bfeta_{0} = \0$, $k = 1$.

    \WHILE{not converge}

      \STATE 1. Compute Riemannian gradient $\bP_{k} = \grad f(\bX_{k})$ by Algorithm~\ref{alg:compute_riemannian_grad}.

      \STATE 2. Compute a conjugate direction with PR+:
             $\bfeta_k = - \bP_{k} + \beta_{k} \mT_{\bX_{k-1} \rightarrow \bX_{k}}( \bfeta_{k-1}) \in T\M_{r}$.

      \STATE 3. Find a step size $t_{k} = \min_{t} f(\bX + t\bfeta_{k}) $

      \STATE 4. Update $\bX_{k+1} = R_{\bX_{k}}(t_{k} \bfeta_{k})$.

      \STATE 5. $k = k + 1$.

    \ENDWHILE

    \ENSURE $\bX \in \M_{r}$.

  \end{algorithmic}
  \caption{LRGeomCG (Algorithm 1 in~\cite{vandereycken2013lowrank})}
  \label{alg:LRGeomCG}
\end{algorithm}
%\end{small}

Particularly, in Step 2 of Algorithm~\ref{alg:LRGeomCG}, $\beta_{k}$ is determined by a Polak-Ribi$\grave{e}$re (PR+) rule~\cite{vandereycken2013lowrank}:
{\small
\begin{eqnarray}\label{eq:beta}
\beta_{k} = \frac{\grad f(\bX_{k})^{\trsp}(\grad f(\bX_{k}) - \grad f(\bX_{k-1}))}{\langle \grad f(\bX_{k-1}),\grad f(\bX_{k-1})\rangle}.
\end{eqnarray}
}
\noindent
The step size $t_{k}$ in Step 3, given a search direction $\bfeta_{k} \in T_{\bX_{k} \M_{r}}$, is chosen such that
{\small
\begin{align}\label{eq:choose_step_size}
  f(R_{\bX_{k}} (t_{k} \bfeta_{k})) \leq f(\bX_{k}) + c_{1} t_{k} \langle \grad f(\bX_{k}), \bfeta_{k} \rangle ,
\end{align}
}
{\small
\begin{align}\label{eq+choose_step_size2}
  | \langle \grad f(R_{\bX_{k}} (t_{k} \bfeta_{k})), & \mT_{\bX_{k-1} \rightarrow \bX_{k}} (\bfeta_{k}) \rangle |  \nonumber \\
                           & \leq c_{2} | \langle \grad f(\bX_{k}), \bfeta_{k} \rangle |  ,
\end{align}
}
\noindent
where $0 \leq c_{1} \leq c_{2} \leq \frac{1}{2}$.
The notation $\mT_{\bX_{k-1} \rightarrow \bX_{k}} (\bfeta_{k-1})$ in Step 2 and $R_{\bX_{k}} (t_k \bfeta_{k})$ denote \emph{vector transport} and \emph{retraction}, respectively.
For more details of above two operators and the geometry of $\M_{r}$, see~\cite{vandereycken2013lowrank} and the references therein.

%To compute $\grad{f(\bX)}$, we need to calculate the gradient of $f(\bX)$ on Euclidean space.
%Firstly, the gradient of $\ell(\bX)$ w.r.t. $\bX$, denoted by $\widehat{\G}$, can be calculated by
%\begin{small}
%\begin{eqnarray}\label{eq:grad_X_Euc}
%\widehat{\G}_{ij} = \frac{\partial \ell(\bX)}{\partial \bX_{ij}} = \sum_{z=1}^{L-1} T_{ij}^{z} \cdot h(T_{ij}^{z} \cdot(\theta_{z} - \bX_{ij})).
%\end{eqnarray}
%\end{small}
%where $ij \in \Omega$. Note that the gradient of $ \Upsilon (\bX)$ w.r.t. $\bX$ is $ \U\diag(\bsigma - \nu/\bsigma^3)\V^{\trsp}$ at $\bX = \U\diag(\bsigma)\V^{\trsp}$. The gradient of $f(\bX)$ w.r.t. $\bX$ in Euclidian space, denoted by $\G$, can be computed by
%\begin{small}
%\begin{eqnarray}
%\G = \widehat{\G} + \lambda \U\diag(\bsigma - \nu/\bsigma^3)\V^{\trsp}.
%\end{eqnarray}
%\end{small}
%Once $\G$ is computed, $\grad{f(\bX)}$ can be calculated according to equation (\ref{eq:grad}). The details of computation can be found in Appendix A.

\subsection{Optimization of The Subproblem of $\bE$}

Similarly, to update $\bE$, we consider the augmented Lagrangian function \wrt. $\bE$ as below:
%\begin{align}\label{eq:subproblem_wrt_E_1}
%  \calL (\bE) = & || \bE ||_{1,2} + || \bL - \bX - \bE + \frac{\blambda}{\mu} ||_F^2 -\frac{1}{2\mu}|| \blambda ||_F^2  \nonumber \\
%              = & || \bE ||_{1,2} + || \bN - \bE ||_F^2 - \frac{1}{2\mu}|| \blambda ||_F^2   ,
%\end{align}
{\small
\begin{align}\label{eq:subproblem_wrt_E_1}
  \calL (\bE) = & || \bE ||_{1,2} + || \bN - \bE ||_F^2 - \frac{1}{2\mu}|| \blambda ||_F^2   ,
\end{align}
}
\noindent
where $\bN = \bL - \bX + \frac{\blambda}{\mu}$.
We thus obtain the subproblem \wrt. $\bE$ as below:
{\small
\begin{align}\label{eq:subproblem_wrt_E_2}
  \min_{\bE} & || \bE ||_{1,2} + || \bN - \bE ||_F^2 .
\end{align}
}
\noindent
The above problem can be efficiently solved by the following column-wise soft-thresholding operator~\cite{xiao2015FaLRR}:
{\small
\begin{align}\label{eq:column_wise_soft_thresholding}
  \bE_{i} = \calS_{\alpha}(\bN_{i}) = \left\{
    \begin{aligned}
      & \zerocolumn,~\text{if}~ ||\bN_{i}||_2 \leq \alpha   \\
      & \bN_{i} - \frac{\alpha \bN_{i}}{|| \bN_{i} ||_2},~\text{otherwise,}
    \end{aligned}
    \right.
\end{align}
}
\noindent
where $\bE_{i}$ and $\bN_{i}$ denote the $i$-th column of $\bE$ and $\bN$,
%and $\alpha = \frac{2}{\mu}$.
and $\alpha = \frac{1}{2}$.

\subsection{Post Process}
As we recover the low rank matrix $\bX$ from the label assignment matrix $\bL$,
it needs a post process to generate the prediction label for each testing data.
Intuitively, there might be several possible ways to select as below:
\begin{itemize}
  \item $[AVE]$ A prediction score matrix $\bX^{*} \in \dsR^{n \times {\calC}}$ is obtained by averaging $[ \bX_1, \bX_2, ..., \bX_{\calC}]$.
		The indicator of the highest value in each row of $\bX^{*}$ is regard as the predicted label for the corresponding data.
  \item $[VOTE]$ For $\bX_j$ where $1 \leq j \leq m$, we set $\bI_{ij}$ as the index of the highest value in the $i$-th row of $\bX_j$.
  		Each $\bI_{ij}$ contributes one point to the $i$-th data with label $\bI_{ij}$.
  		Finally, we choose the label with the most votes to be the prediction results.
  \item $[GATE]$ $\bX_j$ where $1 \leq j \leq m$ are transformed through softmax function as a gate of the corresponding confidence score matrix, which is normalized by $L_2$ norm.
  		As $\bX_j$ contains outlier rejection information, we product the $\bX_j$ by the corresponding score matrix after softmax and normalization.
  		The similar process as the first strategy is applied to the above results, which predictes the final labels.
\end{itemize}
According to the experiments, the last strategy ``GATE'' outperforms than the others.
So We choose the last one as the post process for our fusion method and analyse inner reasons during the experiment section.

\section{Empirical Studies}

In this section, we first visualize the results of our approach on the synthetic dataset.
Then we evaluate the perfomance on eight publicly available real-world data sets,
UCF101~\cite{soomro2012ucf101}, CIFAR-10~\cite{krizhevsky2009learning}, CIFAR-100~\cite{krizhevsky2009learning}, Oxford-IIIT-Pet~\cite{parkhi2012cats}, PASCAL VOC 2007~\cite{pascal-voc-2007}, Oxford Flower 17~\cite{nilsback2006visual}, Pascal Sentence~\cite{rashtchian2010collecting} and Wikipedia~\cite{rasiwasia2010new}.
The strategy analysis for post process is detailed in the last subsection.

%Details on feature extraction for each dataset are described in the second subsection.
%, and the confidence scores are set as outputs of LIBLINEAR.
%Then we put the confidence scores as the input X of our alogrithm.

Five state-of-the-art fusion methods are compared with our proposed alogrithm :
(1)~Average Score Fusion(ASF). we directly average the confidence scores predicted by classifiers. We use $L_2$ norm on each classifier's results for normalization.
(2)~Multiple Kernel Learning (MKL).  MKL learn a weight w for each classifiers, the final score are obtained from function $f(s)=w^{T}s, \sum w = 1$. We use liblinear-mkl\footnote{\url{http://www.csie.ntu.edu.tw/~b95028/software/liblinear-mkl/}} toolbox to train our MKL model.
(3)~Robust Convex Ensemble Clustering (RCEC)\cite{gaoijcai2016robust}. The core alogrithm in RCEC can be modified to recover X from L in~\ref{eq:l12_mf}, and the post process is same as our method.
(4)~Feature Weighting via Optimal Thresholding(FWOT)\cite{xuiccv2013feature}. The author propse a method to learn thresholding, smoothing parameters and weights in a joint framework to combine prediction confidence scores of different features.
(5)~LPBoost~\cite{gehler2009feature}. In this method, a variant linear combination is applied to multiple classifiers to boost performance.

The parameter $\mu$ is selected from $\{0.1, 1, 5\}$ and $\rho$ is select from $\{1.01, 1.05, 1.1\}$ in all our experiments.
Among different parameters, there are small differences of the prediction performance.
LIBLINEAR~\cite{fan2008liblinear} is adopted for our basic classification toolbox.
The parameters are set as default in LIBLINEAR and liblinear-mkl, unless otherwise specified.
For RCEC method, we use 5 folder cross validation to select the best parameter $\beta$ from $\{0.01,1,2,4,6,...,20\}$ and set $\lambda = 0.1, \gamma = 0.01$, which are suggested by author~\cite{yiicdm2012robust}.
For FWOT, the smoothing parameter candidates are $\{0.5, 0.6, ... , 0.9\}$ and the parameter $C$ in is selected from $\{10^{-4},10^{-2},...,10^{4}\}$ according to cross-validation, which is same in~\cite{xuiccv2013feature}.
For LPBoost, we use cross-validation to pick up the best $\nu$ from $\{0.5,0.6,0.7,0.8,0.9\}$ suggested in~\cite{xuiccv2013feature}.

\subsection{Synthetic Experiments}

In synthetic experiments, we set $n = 280$, $m = 15$, $\calC = 20$ to generate $n$ instances, which are randomly labeled with $c_i, 1 \leq i \leq n, 1 \leq c_i \leq \calC$.
And we regard these instances as the ground truth data.
Then we randomly permute $40\%$ instances' labels, repeating $m = 15$ times to get fifteen classifiers' results.
Deriving an indicator matrix $\bL \in \dsR^{280 \times 300}$ from the synthetic classifiers' results as the input of our proposed alogrithm,
the visualization results are illustrated in Figure~\ref{fig:ensemble_cluster}.

Each subfigure contains $n\times m$ small narrow blocks visualized in diverse color according to different label.
The first picture~\ref{subfig:gt_pdf} show the visualized ground truth data, which are copied $m$ times.
And $m$ noised classifiers' results are pictured as~\ref{subfig:er_pdf}.
We recover the low rank matrix $\bX$ from the noised matrix by using our late fusion method, which is visualized as~\ref{subfig:re_pdf}.

\begin{figure}[htp]
\center
    \subfigure[Ground Truth]{
    \label{subfig:gt_pdf}
    \includegraphics[width=0.14\textwidth]{resource/ground_truth.eps}
    }
    \subfigure[Random Noise]{
    \label{subfig:er_pdf}
    \includegraphics[width=0.14\textwidth]{resource/random_error.eps}
    }
    \subfigure[Outlier Reject]{
    \label{subfig:re_pdf}
    \includegraphics[width=0.14\textwidth]{resource/recover.eps}
    }
    \caption{visualized results of the synthetic dataset} \label{fig:ensemble_cluster}
\end{figure}

%% Flower 17 Time [3.14] [229] [0.26] [1.62]

\subsection{Feature Extraction}
We describe the details of the feature extraction for each dataset we use.
All CNN features are extracted using Caffe~\cite{jia2014caffe}, a deep learning framework.

In UCF101 dataset, five different features are extracted from the following:
%\begin{itemize}
%  \item `fc6' of C3D~\cite{tran2015learning}
%  \item `pool5/7x7\_s1' from GoogleNet~\cite{szegedy2015going}
%  \item `pool5' of ResNet-152~\cite{he2015deep}
%  \item two `fc6's of Two-Stream~\cite{simonyan2014two}
%\end{itemize}
`fc6' of C3D~\cite{tran2015learning},
`pool5/7x7\_s1' from GoogleNet~\cite{szegedy2015going},
`pool5' of ResNet-152~\cite{he2015deep},
and two `fc6's of Two-Stream~\cite{simonyan2014two}.
Given a video, we extract the above five features following the same process in~\cite{simonyan2014two,tran2015learning}(GoogleNet and ResNet-152 feature are same as spatial net in~\cite{simonyan2014two}).
All models we used are public available on Internet. %% provided by authors.

In Cifar10 dataset, the features are extracted from the `global\_pool' layer of ResNet-20,32,44,56,110~\cite{he2015deep} and of pre-activation ResNet-20,32,44,56,110,164 models~\cite{he2016identity}.
One feature is extracted from `pool3' layer's features of cifar10\_full model described in caffe example prototxt
\footnote{\url{https://github.com/BVLC/caffe/blob/master/examples/cifar10/cifar10_full.prototxt}}
.
Two modified networks VGG16 and GoogleNet are used to extract final pooling features~(Modify the stride 2 in conv4 and conv5 layers to stride 1 and fc layers to global pooling layers), trained the same in~\cite{he2015deep}
Each ResNet model is trained following the standard procedure in~\cite{he2015deep,he2016identity}, and cifar10\_full model is trained by an official shell
\footnote{\url{https://github.com/BVLC/caffe/blob/master/examples/cifar10/train_full.sh}}
before extracting features.
In Cifar100 dataset, we use `global\_pool' layer features of ResNet-20,32,44,56,68,110 and of pre-activation ResNet-20,32,44,56,110,164 models, following the same training process as in~\cite{he2015deep,he2016identity}.

In all other experiments, eight different features are extracted from the following:
%\begin{itemize}
%  \item `fc6' of VGG16,VGG19~\cite{chatfield2014return}
%  \item `fc6' of AlexNet,CaffeNet~\cite{krizhevsky2012imagenet}
%  \item `pool5/7x7\_s1' from GoogleNet~\cite{szegedy2015going}
%  \item `pool5' of ResNet-50,101,152~\cite{he2015deep}
%\end{itemize}
``fc6'' of VGG16, VGG19~\cite{chatfield2014return},
``fc6'' of AlexNet,CaffeNet~\cite{krizhevsky2012imagenet},
``pool5/7x7\_s1'' from GoogleNet~\cite{szegedy2015going},
and ``pool5'' of ResNet-50,101,152~\cite{he2015deep}.
We use the offical pre-trained model provided by Caffe Model Zoo
\footnote{\url{https://github.com/BVLC/caffe/wiki/Model-Zoo}}
.
For above eight models, we simply extract features with no extra operation.

\subsection{Real-world Experiments}

In this section, we introduce the details about our experiments on each data set,
and demonstrate the empirical results in Table~\ref{table:total_acc} and Table~\ref{table:total_time}.
Five data sets are used to compare the performance on each contrast algorithm, including UCF-101, CIFAR-10, CIFAR-100, Oxford-III-Pet and PASCAL VOC 2007.
And the computational cost of cross-validation is not counted in FWOT and LPBoost.

%\subsubsection{Experiment on UCF 101 dataset}
For action recognition, we present the comparison results on UCF 101 dataset.
UCF101 contains 13320 videos collected from YouTube, categoried into 101 human action classes.
We use the official split-1 described in~\cite{soomro2012ucf101}, which have 9573 training videos and 3783 testing videos.
As mentioned above, we extract fives features and train different classifier on each feature with the default parameters by liblinear.
Then by appling five classifiers on test data, we obtain $5 \times 101$ confidence scores for each video, which forms the input matrix $\bX$ of our algorithm.
Results are shown in Table~\ref{table:total_acc} and \ref{table:total_time}, in which we illustrate the performance of six fusion methods.
To be noted, FWOT is too slow to train, we didn't show its result (more than one day).

\begin{table*}[t]
\small
\centering
\caption{Mean accuracy on real-world data sets.}
\begin{tabular}{c|c|c|c|c|c}
\hline
Method              & UCF-101          & CIFAR-10         & CIFAR-100        & Oxford-IIIT-Pet & PASCAL VOC 2007    \\\hline
ASF                 & 86.78\%          & 94.21\%          & 76.91\%          & 92.46\%         &   90.51\%          \\
MKL                 & 85.38\%          & 94.15\%          & 72.71\%          & 87.13\%         &   85.82\%          \\
RCEC                & 86.60\%          & 94.99\%          & 77.13\%          & 92.11\%         &   90.92\%          \\
FWOT                & -                & -                & -                & 92.90\%         &   91.20\%          \\
LPBoost             & 86.24\%          & 94.88\%          & 76.59\%          & 92.73\%         &   91.42\%          \\\hline
\textbf{Ours}       & \textbf{89.03\%} & \textbf{95.11\%} & \textbf{77.72\%} & \textbf{92.90\%}&   90.98\%          \\
\hline
\end{tabular}
\begin{tablenotes}
\footnotesize
\item[*]
  FWOT costs excessive time for training UCF-101, CIFAR-10 , CIFAR-100~(over 24 hours), so we omit these results indicated by ``-''.
\end{tablenotes}
\label{table:total_acc}
\end{table*}

\begin{table*}[t]
\small
\centering
\caption{Experimental results on real-world data sets. Computational time is recorded in seconds.}
\begin{tabular}{c|c|c|c|c|c}
\hline
Method              & UCF-101    & CIFAR-10  & CIFAR-100  & Oxford-IIIT-Pet & PASCAL VOC 2007 \\\hline
RCEC                & 86.60      & 81.41     &  255.51    & 18.82           &   16.45         \\
FWOT                & $\infty$   & $\infty$  &  $\infty$     & 61697           &   19674         \\
LPBoost             & 110.81     & 72.45     &  4750.9    & 43.67           &   7.46          \\\hline
Ours                & 421.23     & 5.384     &  417.26    & 76.16           &   16.21         \\
\hline
\end{tabular}
\begin{tablenotes}
\footnotesize
\item[*]
  FWOT costs excessive time for training UCF-101, CIFAR-10 , CIFAR-100~(over 24 hours), so we omit these results indicated by ``$\infty$''.
\end{tablenotes}
\label{table:total_time}
\end{table*}

%\subsubsection{Experiment on Cifar10 dataset}

In CIFAR-10 dataset, totally 60000 32x32 colour pictures are categoried into 10 classes, with 6000 pictures per class.
There are 50000 training data and 10000 testing data.
Fourteen CNN features are extracted from multiple state of the art models.
This experiments exhibit the ability of our algorithm to handle large scale data.

As we have 50000 training data and 10000 testing data,
FWOT algorithm has one step to deal with $n\times n$ matrixes, which are excessive to train~(more than one day).
LPBoost algorithm needs to create a $(n\calC - n) \times (1 + n + m)$ constraint matrix linear programming solvers
\footnote{\cite{gehler2009feature} use the MOSEK interior-point solver, see www.mosek.com.},
which occupies too much memory and still consumes long time to train, so we solve this problem by using sparse matrix storage and parallel during matlab.
We also use `-s 2' parameters in liblinear to speedup training process.
Results of total six contrast algorithms are shown in Table~\ref{table:total_acc} and \ref{table:total_time}.

%\subsubsection{Experiment on Cifar100 dataset}

CIFAR-100 is similar with CIFAR-10, except it has 100 classes containing 600 images each.
There are 50000 training data and 10000 testing data as CIFAR-10.
Twelve CNN features used in this experiment are described in the feature selection section.
Training and comparison process are the same as CIFAR-10.
Detailed results are shown in Table~\ref{table:total_acc} and \ref{table:total_time}.

%\subsubsection{Experiment on Oxford-IIIT-Pet dataset}

The Oxford-IIIT-Pet dataset contains 7349 images covering 37 different breeds of cats and dogs, 2371 images for cats and 4978 images for dogs.
There are roughly 200 pictures in each category with a large variations in multiple views, such as pose, scale and lighting.
We randomly sample fifty percent pictures as our training data and the others as testing data.
And then we train svm on each single feature and fuse the predictions by our proposed method.
We repeat the sample , training and fusion process five times to get fair results.
The comparison results are shown in Table~\ref{table:total_acc} and \ref{table:total_time}.

%% VOC07
PASCAL VOC 2007 is a dataset for object detection and image classification.
In our experiments, we only use the image level annotation information.
Since each image may contain more than one label, we filter the multi-label images to obtain 6146 clean data.
There is similar procesure of sampling, classification and fusion as in Oxford-IIIT-Pet.

\subsection{Post Strategy Selection}

In this section, we compare three proposed post strategies on three small datasets: Wikipedia~\cite{rasiwasia2010new}, Oxford Flowers 17~\cite{nilsback2006visual} and Pascal Sentence~\cite{Li2006One}.
Wikipedia dataset contains paragraphs of text for each picture, and total 2866 image-text pairs are categoried into 10 class.
Only the image label annotation is used in our experiments.
PASCAL VOC 2007 contains 9963 images annotated with object-level information, which means each picture may have zero or more than one class.
We choose 6146 images classified into 20 classes by filtering out the no-single class data.
Oxford Flowers 17 has 17 flower categories with 80 images for each class.
Pascal Sentence contains 1000 images annotated by caption information and categoried into twenty classes.
Fifty percent of data are sampled as training set and the others as testing set, and experiments are repeated five times on above three datasets.

The $[$AVE$]$ strategy outperforms than the $[$VOTE$]$ in our experiments, which may be caused by the soft average method reserves more detail information than the hard voting method.
In all our experiments, the accuracy performance of the $[$GATE$]$ strategy is higher than the others.
Since the recover matrix $\bX$ contains the outlier rejection information,
it is more robust to take advantage of the recovered results as a confidence gate operating on the original confidence score matrix.

\begin{table}[ht]
%\small
\centering
\caption{Mean Accuracy with Different Strategies}
\begin{tabular}{c|c|c|c}
\hline
Strategy            & Wikipedia         & Flowers 17          & Pascal Sentence     \\\hline
[AVE]               & 47.78\%           & \textbf{96.77\%}    & 66.40\%             \\\hline
[VOTE]              & 47.85\%           & 96.61\%             & 66.20\%             \\\hline
[GATE]              & \textbf{49.03\%}  & \textbf{96.77\%}    & \textbf{67.20\%}    \\
\hline
\end{tabular}
\label{table:strategy}
\end{table}

\section{Conclusion}

To relieve the performance degeneration in late label fusion from multiple sources, we formulate this problem as matrix factorization which recovers the underlying low-rank matrix and preserves the consistency among different sources.
%To deal with the dispute on combining multi-view predictions, we formulate the late fusion problem as a matrix recover problem.
%Our proposed algorithm is based on matrix factorization, which is unclear to consider smooth loss function from earlier research.
Existing matrix factorization algorithms usually consider smooth loss, while it is unclear how to extend them to handle nonsmooth robust loss, \eg, $\ell_{1,2}$ loss.
We propose an ALM based algorithm to solve matrix factorization problems with nonsmooth loss.
%Specifically, our proposed algorithm first rejects outliers on a label indicator matrix by $\ell_{1,2}$ loss,
%and then generates the predictions by combining robust outlier rejection information and prediction confidence scores from basic classifiers.
Empirical studies on large scale real-world data sets demonstrate that our proposed method can effectively improve performance compared to the state-of-the-art late fusion methods.

\begin{quote}
\begin{small}
  \bibliographystyle{aaai}
  \bibliography{RLF_MF}
\end{small}
\end{quote}

\end{document}