### 2017-06-04 Rebuttal_ICCV.txt
final version of rebuttal

We thank all the reviewers for the comments and suggestions. We take the comments seriously and will carefully revise the paper.

To Reviewer 1:

Thanks for the comments.

- While you generally recognized our technical merit, your major concern is that we did not compare our method to more other fusion methods in which fusion is implicitly tackled at the training stage. Following your suggestion, we conducted more experiment and compared to the Bayesian-based method [1] suggested by you. Because [1] is not directly applicable to fusion, some necessary modification is made. A limitation of [1] is that it requires the same dimensionality of features for fusing, which is not true in many real world applications. The global_pooling layer always generates 64-dimension features. We thus choose the global_pooling layer's feature of all residual-based network used in Table 4 as the input of the joint Bayesian method [1]. In this way, we have feature dimension = 64, 50000 instances (10 models) for total 10 classes. The accuracy of Joint Bayesian [1] is 93.8. Using the SVM predictions with the global pooling feature as input, the accuracy of our method is 95.0, which outperforms the method suggested by you. Additionally, because our method does not require the input feature to be of the same dimension, our method can be applied to a broader range of applications.

- Other two minor issues:
(i)In terms of efficiency, our method and RCEC are the two most efficient methods (See Table 2). The two algorithms have a similar range of running time. On the other hand, our method outperforms RCEC in accuracy for all the five datasets (See Table 2).

(ii) Our method and ResNeXt are two different directions for image classification. Our method can further take ResNeXt as additional input. We conducted the experiments by taking ResNeXt as another input, and our method outperforms ResNeXt by about 3%, which is significant and worthwhile.

[1] Bayesian face revisited: A joint formulation. In ECCV, 2012
We will cite [1] and discuss more implicit fusion methods in our revised version.

To  Reviewer 2:

Thanks for your comments that the paper proposes a new way for information fusion. We also appreciate your suggestions for improving the paper.

Your major concern is that we should add three more experiments. 
To address the concerns, we conducted all the suggested experiments and report the result on CIFAR-10 in this rebuttal due to space limit.

1. SVM-prediction V.S. CNN-prediction as the input of late fusion.
We used SVM prediction as input in our original submission because some of the compared algorithms cannot take CNN predictions as input (see L801 to L803 for our justifications). Nevertheless, we additionally used CNN-predictions learned by end-to-end CNN models as the input of late fusion to address your comments. The accuracy of our method by fusing the CNN-predictions of each model used in Table 1 is 95.93, higher than the fusion of SVM-predictions (95.11). The only two compared methods that are able to use CNN prediction as input are RCEC and average fusion. The accuracies of RCEC and average fusion using CNN prediction are 95.36 and 94.97. Again, our method outperforms the compared methods.

2. Neural network based fusion V.S. Our method

Note that directly training a fusion model that fuses several entire CNN models including all convolutional layers may be intractable due to excessive GPU memory consumption. We thus perform the following two alternative experiments to address your comment.

---(1) Feature fusion via CNN
We conducted an experiment in which fusion is performed at the last pooling layer of all CNN models. Two strategies were used in our experiment. In the first setting, the features are flattened and concatenated (we denote this strategy by 'concatenation'). In the second setting,  we additionally apply max pooling crossing channel (denoted by 'max pooling'), followed by a fc layer with 10 num_outputs and a softmax layer.  
The results (in accuracy) are as follows:
| concatenation | max pooling | ours |
| --- 95.21 --- | -- 94.93 -- | 95.93|

---(2) Late fusion via CNN
We additionally conducted the experiment in which the fusion of different CNNs is performed at the fully connected layer. Below are the experiment results.
| CNN weighting | average | ours |
| --- 95.26 --- |- 95.12 -| 95.93|

As demonstrated by this comparison, our method still achieves the best performance across all baselines.

3. The impact of the number of instances:
You have also asked the impact of the number of instances. We additionally perform an experiment by setting the different number of instances (number of rows in matrix L). In this experiment, SVM-predictions are used as fusion input. The experiment results (in accuracy) are listed below:
| 1000  |  2000  | 4000  |  6000   |  8000  | 10000 |
| 94.94 |  95.10 | 95.12  |  95.10   |  95.11  | 95.11   |

We can observe that our method achieves stable performance when the number is greater than 2000. We will include all new experimental results into the revised paper.

For Reviewer 3:

Thanks for your kind advice. The caption will be revised.
The C represents the number of classes. There are usually not many classes in most real-world applications. In this situation, the complexity is acceptable.
If C is quite large for some specific problems (e.g. YouTube-8M includes 4716 classes). A trade-off method may be splitting all classes into multiple parts, performing late fusion on each part and finally merging together. The above discussions will be included in our revision.