### 2018-07-07 BACKUP/cvpr-2017/Fusion-cvpr2017-RebuttalKit/rebuttal-fusion.txt
update for minor

\section{Overall}
We thank all the reviewers.
R3 is positive.
R1 has some suggestions about technical analysis.
R2 ...
We take the comments seriously and will carefully revise the paper. 
Typos will be corrected. Code will be released.

\section{For Assigned Reviewer 1}

Q1: The technical part in Sec 3.2 a little over-sold.

A1:
First, ALM is usually used for convex problem~[50] and (G. Liu etal, TPAMI 2013),
but our Problem~(2) is non-convex.
Therefore, the convergence behavior of Algorithm 1 may be unclear.
Second, to make the paper self-contained, it is necessary to present every phase of Algorithm~1.
Third, we explicitly indicate the connection between the definition (Eq (8)) and the computation (Algorithm 2) of Riemannian gradient by presenting Lemma 1.
This did not present in the previous work.
To make the contributions more clear, we will provide more references and explain the difference between our paper and previous works as suggested.

\section{For Assigned Reviewer 2}

Q1. Weakness of technical novelty and significance}

A1:
Previous works~[50] that formulate late fusion as matrix completion are based on nuclear norm.
They cannot handle large-scale problems.
We thus propose an efficient matrix factorization-based algorithm.
On one hand, previous matrix factorization algorithms only consider smooth loss rather than non-smooth robust loss.
On the other hand, previous ALM-related papers consider convex problems, so it is unclear how its convergence behavior is in the non-convex case.
To address both issues, this paper proposes a new matrix factorization-based algorithm to handle non-smooth loss and provides convergence analysis.

Q2. Problem~(6) can be solved by SVD or alternating minimization.

A2:
In P. Jain et al., 2013, the authors provide global optimality of $\epsilon$-solution (i.e. the loss function satisfies $|| M - X ||_F^2 \leq \epsilon$. See Theorem 2.5 in this paper.) of alternating minimization, rather than a strictly unique global optimum.
Our paper provides the convergence guarantee of a strictly local minimum where the gradient w.r.t. $\bX$ vanishes.
Thus, their result and ours may not be comparable.

Q3. Algorithms 2 and 3 may be unnecessary.}

A3: 
The algorithm should be self-contained and it may be necessary to describe solutions for any formula.
But we are pleasure to simplify our presentation.

Q4. Will the sequence of $\lambda_k$ be bounded as required, or that accumulation points even exist?

A4:
Boundedness of $\blambda_k$:
After the inner loop (Step 1,2) of Algorithm 1 converges, we have 
$\calL(\bX_{k+1}, \bE_{k+1}, \blambda_{k}, \mu_{k}) = ||\bE_{k+1}||_{1,2} + \frac{1}{2\mu_{k}}(||\blambda_{k+1}||_{F}^2 - ||\blambda_{k}||_{F}^2)$.
Due to the convergence and $\frac{1}{\mu_{k}} \rightarrow 0$ when $k \rightarrow + \infty$, we have $||\blambda_{k+1}||_{F}^2 - ||\blambda_{k}||_{F}^2 < +\infty$, and thus the boundedness of $\blambda_{k}$.

As proof of Theorem 1, $k \rightarrow \infty$, $\bL - \bX_{k} - \bE_{k}$ and the gradient of $\calL$ w.r.t. $\bX_{k}$ vanishes.
Thus $\bX_{k}$ is a stationary point as long as $\bX_{k}$ is generated according to Algorithm 1.

Q5. Empirical issues in Table 1 and Table 5.

A5:
As shown in Table 1, our method outperforms LPBoost significantly on UCF-101 and CIFAR-100,
and our fusion method performs well on most datasets comparing with other methods.

Q6. Why not compare against other SOTA fusion methods in Table.5, but just single model results ?

A6:
We compare our method with state-of-the-art single models in Sec 4.5 and list the test error in Table 5 to verify the improvement of our late fusion algorithm.
Test error of competing models are referred from the original papers.
Comparison between our method and other late fusion methods is performed specifically in Sec 4.3 (Table 1 and Table 2).

Q7. Reproducibility - It is obvious that a fused model is better than the best single model.

A7:
It can't guarantee that a fused model can definitely perform better than the best single model.
By comparing with single models, we aim to verify the effectiveness of our method.
The best single model performance on CIFAR-10 is currently in~\cite{xie2016aggregated} (test error 3.58).
Our method fuse multiple single models which all perform worse than~~\cite{xie2016aggregated}, but can achieve low test error on CIFAR-10 (3.47 in Table 5).
This demonstrates the benefit of a good late fusion method.

\section{For Assigned Reviewer 3}

Q1. Discuss how the $\ell_{1,2}$ norm helps with outliers; show what happens if you use $l_2$ norm as usual}

A1:
As indicated in lines 106-110, $\ell_{1,2}$ loss preserves the fidelity within columns of $\bL$ and filters out those columns with low fidelity.
$\ell_{2}$ loss is known to be sensitive to outliers.
$\ell_{2}$ loss is not robust to large outliers and the result may be biased to them.
We will add a $\ell_{2}$ variant of our approach as a baseline in the revised version to illustrate the difference.

Q2. Discussions about classifiers are adversarial.

A2:
In fact, Sec 4.1 and Figure 2 show a similar situation: 40\% predicted labels are switched from ground-truth (adversarial classifiers).
In this situation, our algorithm performs well and recovers most of the results.

Q3. Add classifiers that predict uninterested things.}

A3:
Thanks for this interesting and helpful suggestion.
We will investigate this problem in our future work.